### 2024-09-12 parts/correction.tex
enormous rewrite

\AP We will discuss at the end of \cref{sec:poly-to-n-poly}
why the
example of \cref{thm:counter-example} is minimal in the number of
indeterminates, which first requires us to provide a \emph{correct} analogue of
\cref{karh:thm}. The
counterexample provided by \cref{def:bad-polynomial}
relies on the fact that
$\CoveredPoly$ is not stable under fixing indeterminates, while
\kl{$\Nat$-rational polynomials} are. Indeed, the polynomial $\BadPoly$
introduced in \cref{def:bad-polynomial}
satisfies $\BadPoly(X,Y,1) = 3X^2 +
3Y^2 - 2XY$, which has a negative coefficient for a \kl{maximal monomial},
namely $-2XY$. In this section, we prove that closing $\CoveredPoly$ under
variable assignments is enough to recover from
\cref{karh:thm}. We
use the following notation to fix the value of some indeterminate, if $P(X,Y)$
is a polynomial in $\Rel[X,Y]$, then $\intro*\restr{P(X,Y)}{X = 1}$ is the
polynomial $P(1,Y) \in \Rel[Y]$. More generally, if $\nu$ is a partial function
from $\vec{X}$ to $\Nat$, written $\nu \colon \vec{X} \topartial \Nat$, the
restriction $\restr{P(\vec{X})}{\nu}$ is the polynomial with indeterminates
$\vec{Y} \defined \vec{X} - \dom(\nu)$ obtained by fixing the variables of the
domain of $\nu$.

@@

\begin{theorem}[restate=corrected-version:thm,label=corrected-version:thm]
	Let $P \in \Rel[\vec{X}]$.
	The following are equivalent:
	\begin{enumerate}
		\item \label{corrected-1:item} $P \in \CorrectPoly$,
		\item \label{corrected-2:item} $P$ is \kl{represented} by a \kl{$\Nat$-rational series},
		\item \label{corrected-3:item} $P$ is \kl{represented} by a \kl{$\Nat$-polyregular function},
		\item \label{corrected-4:item} $P$ is \kl{represented} by a \kl{star-free $\Nat$-polyregular function},
	\end{enumerate}
	Furthermore, the properties are decidable, and conversions effective.
\end{theorem}

@@

\begin{remark}
    \label{undecidable-non-negative:lem}
    \proofref{undecidable-non-negative:lem}
    Checking whether a polynomial
    $P \in \Rel[\vec{X}]$ is \kl{non-negative} is undecidable.
\end{remark}

The proof of \cref{corrected-version:thm} is divided into two parts. First, we
provide in \cref{sec:n-poly-to-poly} a fine combinatorial understanding of what
functions can be computed in $\NPoly$ and $\ZPoly$. This allows us to prove
that \kl{$\Nat$-polyregular functions} are in $\CorrectPoly$
(\cref{n-rat-correct:lem}). Then, in \cref{sec:poly-to-n-poly} we will show how
to compute polynomials in $\CorrectPoly$ using $\NPoly$
(\cref{lem:correct-to-n-rat}). With this in hand, we will next generalize these
results to polynomials in $\Rat[\vec{X}]$ in \cref{sec:rel-to-rat}.

@@

\AP To address our first goal, let us define that a function $q$ is a
\intro{(noncommutative output) polynomial function} from $\Nat^p$ to $\Sigma^*$
whenever there exists words $\alpha_0, \dots, \alpha_p \in \Sigma^*$, and words
$u_1, \dots, u_p \in \Sigma^*$, such that $q(X_1, \dots, X_p) = \alpha_0
	\prod_{i = 1}^p u_i^{X_i} \alpha_i$. That is, $q$ is syntactically defined by a
non-commutative monomial over the monoid $\Sigma^*$. Almost by
definition, \kl{polynomial functions} are \kl{commutative} \kl{polyregular
	functions}.

@@

Let us highlight that the multiplicative factor $\omega$ is necessary in
\cref{n-poly-combinatorics:lem}. Indeed, the function $f \colon \set{a}^* \to
	\Nat$ defined as $0$ when the input is of odd length and $1$ when the input is
of even length is \kl{$\Nat$-polyregular}, but $f(a^X)$ is not a polynomial. We
can trade off this multiplicative factor for a constant term addition under the
extra assumption that the function is \kl{star-free polyregular}, as described
in the following \cref{n-sf-combinatorics:lem}.

@@

\begin{corollary}
	\label{n-rat-correct:lem}
	Let $P \in \Rel[X_1, \dots, X_p]$ be an \kl{$\Nat$-rational polynomial}.
	Then,
	$P \in \CorrectPoly$.
\end{corollary}
\begin{proof}
	Let $f$ be the \kl{commutative}
	\kl{$\Nat$-polyregular function}
	with domain defined as $\Sigma \defined \set{a_1, \dots, a_p}$
	that \kl{represents} $P$.
	Using \cref{n-poly-combinatorics:lem},
	there exists a number $\omega \in \Nat_{\geq 1}$
	and \kl{natural binomial function} $Q$
	such that
	for all $n_1, \dots, n_p \geq 1$:
	\begin{equation*}
		f\left(
		\prod_{i = 1}^p a_i^{\omega n_i}
		\right)
		= Q(n_1, \dots, n_p)
		= P(\omega n_1, \dots, \omega n_p)
		\quad .
	\end{equation*}
	Let us remark that for large enough values of $X$, the \kl{binomial function}
	$\binom{X - p}{k}$ coincides with a polynomial whose dominant coefficient
	is $1/k!$ which is \kl{non-negative}.
	We conclude that
	the \kl{maximal monomials} of
	$P(\omega X_1, \dots, \omega X_p)$ are \kl{non-negative},
	and since $\omega \geq 1$, we conclude that
	the \kl{maximal monomials} of $P$ have \kl{non-negative} coefficients.

	Remark that for every partial valuation $\nu \colon \vec{X} \topartial \Nat$,
	the polynomial $\restr{P}{\nu}$ continues to be represented
	by a \kl{$\Nat$-polyregular function}, namely
	$f$ partially applied to a word $u$ and using a restricted input alphabet. As a consequence,
	the \kl{maximal monomials} of
	$\restr{P}{\nu}$ are also \kl{non-negative},
	and
	we have proven that $P \in \CorrectPoly$.
\end{proof}

@@

\AP We have proven that every \kl{$\Nat$-rational polynomial} is in
$\CorrectPoly$, and the rest of this section is devoted to proving the converse
implication. The core of the upcoming \cref{lem:correct-to-n-rat}
leverages a notion of \intro{discrete derivative} which is built by
translating the domain of the polynomial. To that end, let us write
$\intro*\translate{K}$ for the \intro{translation function} that maps a
polynomial $P \in \Rel[X_1, \dots, X_n]$ to the polynomial $P(X_1 + K, \dots,
X_n + K)$.

@@

\AP The key lemma of this section is \cref{lem:correct-to-n-rat}, which is
proved by induction on the number of indeterminates of a given polynomial $P$.
Notice that in \cref{lem:correct-to-n-rat}, polynomials can be represented
using \kl{star-free $\Nat$-polyregular functions}, which is a stronger
statement than the one in \cref{karh:thm}, but is in line with the idea that
polynomials are not periodic functions (this will be made precise in
\cref{zsf-nsf:conjecture}). In order to prove that result, we use the
combinatorial \cref{derivation-stabilises-correct:lem} that allows us to
transform a polynomial $P \in \CorrectPoly$ into a polynomial in
$\Nat[\vec{X}]$ through a well-chosen translation of the indeterminates.

@@

While \cref{lem:correct-to-n-rat} provides an effective conversion procedure,
it does not explicitly state that the membership is decidable to keep the proof
clearer. A similar proof scheme can be followed to conclude that membership is
decidable, and even show that elements in $\CorrectPoly$ are, up to suitable
translations, polynomials in $\Nat[\vec{X}]$. Beware that partial applications
are still needed in this characterization, as \cref{bad-poly-translate:ex}
illustrates. Effectiveness was the last missing piece of our main result
\cref{corrected-version:thm}, which corrects and generalizes \cref{karh:thm}.
\begin{example}
	\label{bad-poly-translate:ex}
	The polynomial $\BadPoly$ is not a
	\kl{$\Nat$-rational polynomial},
	but is \kl{non-negative} and satisfies
	$\translate{10}(\BadPoly) \in \Nat[\vec{X}]$.
\end{example}

\begin{lemma}
	\label{derivation-translation:lem}
	\proofref{derivation-translation:lem}
	Let $P \in \Rel[\vec{X}]$.
	There exists a computable number $K \in \Nat$
	such that the following are equivalent:
	\begin{enumerate}
		\item \label{d-t-correct:item} $P \in \CorrectPoly$,
		\item \label{d-t-transl:item}
		      for
		      all partial functions $\nu \colon \vec{X} \topartial \Nat$,
		      $\translate{K}(\restr{P}{\nu}) \in \Nat[\vec{X}]$,
		\item \label{d-t-transl-fin:item}
		      for all partial functions
		      $\nu \colon \vec{X} \topartial \set{0, \dots, K}$,
		      $\translate{K}(\restr{P}{\nu}) \in \Nat[\vec{X}]$.
	\end{enumerate}
	In particular, the above properties are decidable.
\end{lemma}

@@

\AP Let us complete our analysis of polynomials \kl{represented} by $\NPoly$ or
$\ZPoly$ by considering polynomials with coefficients in $\Rat$, and justify
that all the combinatorial work has already happened in $\Rel$ and $\Nat$. A
first observation is that from \cref{n-sf-combinatorics:lem}, we
know that the polynomials that can be computed by \kl{star-free
$\Nat$-polyregular functions} are going to be suitable combinations of
\kl{natural binomial functions}. This motivates the following definition of the
``polynomial counterpart" of a binomial coefficient, given two numbers $\ell,k
\in \Nat$, $\intro*\pbinom{X - \ell}{k}$ defined as $(X - \ell) \cdots (X -
\ell - k) / k!$,\footnote{ In particular, $\pbinom{X - \ell}{k}$ is defined to
be $1$ when $k = 0$, and $X - \ell$ when $k = 1$. } that we call a
\intro{binomial monomial}, and introduce \intro{natural binomial polynomials}
as $\Nat$-linear combinations of products of \kl{binomial monomials}, i.e., of
the shape: $P(X_1, \dots, X_k) = \sum_{i = 1}^n n_i \prod_{j = 1}^k \pbinom{X_j
- p_{i,j}}{k_{i,j}}$. Similarly, we introduce the class of \intro{integer
binomial polynomials}, which are obtained by $\Rel$-linear combinations of
products of \kl{binomial monomials}.

\AP
These definitions are motivated by the classical result of Pólya that
characterizes polynomials $P$ in $\Rat[X]$ that \intro{integer-valued} (i.e., are such
that $P(\Rel) \subseteq \Rel$) as \kl{integer binomial polynomials}
\cite{POLYA1915,CACHA1996}, which are often called \emph{Newton Polynomials}. Note
that this result straightforwardly extends to multiple indeterminates as we
prove in \cref{integer-binomial-polynomial:lem}.

@@

\begin{theorem}
	\label{integer-binomial-polynomials:cor}
	Let $P \in \Rat[\vec{X}]$, and let $\alpha$ be the minimal number in $\Nat$ such that
	$\alpha P \in \Rel[\vec{X}]$. Then, the following properties are equivalent:
	\begin{enumerate}
		\item \label{int-bin-0:item} $P$ is an \kl{integer binomial polynomial},
		\item \label{int-bin-1:item} $P$ \kl{represented} by a \kl{star-free $\Rel$-polyregular function},
		\item \label{int-bin-2:item} $P$ \kl{represented} by a \kl{$\Rel$-polyregular function},
        \item \label{int-bin-3:item} $P$ is \kl{integer-valued}.
	\end{enumerate}
	These properties are furthermore decidable.
\end{theorem}
\begin{proof}
	Let us first note that implications \cref{int-bin-1:item}
	$\implies$ \cref{int-bin-2:item} $\implies$ \cref{int-bin-3:item} are obvious.
	Now, \cref{int-bin-3:item} $\implies$ \cref{int-bin-0:item} is a direct
	consequence of \cref{integer-binomial-polynomial:lem}.
	Finally, \cref{int-bin-0:item} $\implies$ \cref{int-bin-1:item} follows from the fact
	that $\pbinom{X - p}{k}$ is \kl{represented} by a \kl{star-free $\Rel$-polyregular function}
	defined by hardcoding the output values (in $\Rel$) when $0 \leq X \leq p$, and
	using a \kl{star-free $\Nat$-polyregular function} when $X > p$
	(\cref{binomial-function-star-free:lem}).
	Because
	$\ZSF$ is closed under products and $\Rel$-linear combinations, we conclude.
\end{proof}

\AP Obtaining an analogue of \cref{integer-binomial-polynomials:cor} for
\kl{$\Nat$-polyregular functions} requires a bit more work, as polynomials in
$\Rat[\vec{X}]$ that are \kl{represented} by $\NPoly$ are not exactly
\kl{natural binomial polynomials},
(\cref{natural-binomial-polynomial-positive-bad:ex}). To address the issues
raised by the former example, we introduce the notion of \intro{strongly
natural binomial polynomials}, as the polynomials $P \in \Rat[X]$ such that for
all partial valuation $\nu \colon \topartial \Nat$, $\restr{P}{\nu}$ is a
\kl{natural binomial polynomial}, which characterizes the class of polynomials
that are \kl{represented} by $\NPoly$ (\cref{decide-rat-poly-npoly:cor}).

\begin{example}[restate=natural-binomial-polynomial-positive-bad:ex,label=natural-binomial-polynomial-positive-bad:ex]
	\proofref{natural-binomial-polynomial-positive-bad:ex}
	The polynomial
	$Q(X,Y,Z) \defined \pbinom{X - 4}{1} \pbinom{Y}{1} \pbinom{Z}{1}
		+ 8 \pbinom{Y}{2} + 8 \pbinom{Z}{2} + 4$
	is a \kl{non-negative} \kl{natural binomial polynomial},
	but cannot be computed by a \kl{star-free $\Nat$-polyregular function}.
\end{example}

\begin{theorem}
	\label{decide-rat-poly-npoly:cor}
	Let $P \in \Rat[\vec{X}]$ be a polynomial with \emph{rational}
	coefficients and let $\alpha$ be the smallest number in $\Nat_{\geq 1}$
	such that $\alpha P \in \Rel[\vec{X}]$. Then, 
    the following are equivalent:
	\begin{enumerate}
		\item \label{rat-npoly-1:item} $P$ is \kl{represented} by a \kl{$\Nat$-polyregular function},
		\item \label{rat-npoly-2:item} $P$ is \kl{represented} by a \kl{star-free $\Nat$-polyregular function},
		\item \label{rat-npoly-3:item}
		      $P$ is a \kl{strongly natural binomial polynomial},
        \item \label{rat-npoly-4:item} $\alpha P \in \CorrectPoly$ 
            and $P$ is \kl{integer-valued}.
	\end{enumerate}
	In particular, the properties are decidable.
\end{theorem}
\begin{proof}
    For the implication \cref{rat-npoly-1:item} $\implies$
    \cref{rat-npoly-4:item}, we obtain $\alpha P \in \CorrectPoly$ via
    \cref{corrected-version:thm} by remarking that the \kl{star-free
    $\Nat$-polyregular functions} have output in $\Nat$ and are closed under
    multiplication by a constant $\alpha \in \Nat$. The fact that $P$ is
    \kl{integer-valued} follows from
    \cref{integer-binomial-polynomials:cor} and the fact that
    $\NSF \subseteq \ZPoly$.

    Then, \cref{rat-npoly-4:item} $\implies$
    \cref{rat-npoly-3:item}, because if $\alpha P \in \CorrectPoly$,
    it is represented by a \kl{star-free $\Nat$-polyregular function} $f$ by
    \cref{corrected-version:thm}. Let us now use \cref{n-sf-combinatorics:lem}
    to obtain a number $s \geq 1$ (that only depends on $P$) such
    that for all \kl{polynomial function} $q \colon \Nat^k \to \Sigma^*$, $g
    \circ q(X_1 + s, \dots, X_k + s)$ is a \kl{natural binomial function} $F$
    over $\Nat^k$. Let us write the
    \kl{natural binomial function} $F$ as follows $F(X_1, \dots, X_k) =
    \sum_{\card{\alpha} \leq d_2} e_{\alpha} \prod_{i = 1}^k \binom{X -
    p_{\alpha,i}}{\alpha(i)}$, with $e_{\alpha} \in \Nat$ for all
    $\card{\alpha} \leq d_2$.
    Now, because of the equality 
    $\binom{X - p - 1}{k} + \binom{X - p - 1}{k - 1} = \binom{X - p}{k}$,
    we conclude that we can find a $K \in \Nat$ and coefficients
    $f_{\alpha} \in \Nat$ such that 
    $F(X_1, \dots, X_k) =
    \sum_{\card{\alpha} \leq d_2} f_{\alpha} \prod_{i = 1}^k \binom{X_i - K}{\alpha(i)}$.
    Now, remark that $P$ is an \kl{integer-valued} polynomial of
    degree $d$, and therefore one can write $P = \sum_{\card{\alpha} \leq d}
    c_{\alpha} \prod_{i = 1}^k \pbinom{X_i - K}{\alpha(i)}$ with $c_{\alpha}
    \in \Rel$ for all $\card{\alpha} \leq d$
    (\cref{integer-binomial-polynomial:lem}). 
    As a consequence,
    $\lambda P = \sum_{\card{\alpha} \leq d} \lambda c_{\alpha} \prod_{i = 1}^k \pbinom{X_i - K - s}{\alpha(i)}$,
    and 
    by identifying coefficients in the previous two expressions, we obtain
    $c_{\alpha} = \lambda f_{\alpha}$ for all $\card{\alpha} \leq d$.
    Finally,
    $\lambda P(x_1 + s, \dots, x_k + s) = F(x_1, \dots, x_k)$ for all $\vec{x} \in \Nat^k$,
    and therefore
    $\lambda P (X_1 + s, \dots, X_k + s) =
    \sum_{\card{\alpha} \leq d_2} f_{\alpha} \prod_{i = 1}^k \pbinom{X_i - K}{\alpha(i)}$.
    In particular,
    we get
    \begin{equation*}
        \sum_{\card{\alpha} \leq d} \lambda c_{\alpha} \prod_{i = 1}^k \pbinom{X_i - K}{\alpha(i)} =
        \sum_{\card{\alpha} \leq d_2} f_{\alpha} \prod_{i = 1}^k \pbinom{X_i - K}{\alpha(i)}
    \end{equation*}
    And we conclude by term identification
    by term identification
    that $d = d_2$, and for all $\card{\alpha} \leq d$, $\lambda c_{\alpha} = f_{\alpha}$.
    As a consequence, $c_{\alpha} \in \Nat$ for all $\card{\alpha} \leq d$,
    and $P$
    is a \kl{natural binomial polynomial}.
    Remark that this reasoning remains true whenever we partially apply $P$ to some values
    in $\Nat$, as it amounts to 
    partially applying the function $f$. We have proven that 
    $P$ is a \kl{strongly natural binomial polynomial}.

>>>>

\AP We will discuss at the end of \cref{sec:poly-to-n-poly} why
\cref{thm:counter-example} is minimal in the number of indeterminates, which
first requires us to provide a \emph{correct} analogue of \cref{karh:thm}. Our
counterexample relies on the fact that $\CoveredPoly$ is not stable under
fixing indeterminates, while \kl{$\Nat$-rational polynomials} are. Indeed, the
polynomial $\BadPoly$ satisfies
$\BadPoly(X,Y,1) = 3X^2 + 3Y^2 - 2XY$, which has a negative coefficient for a
\kl{maximal monomial}. Let us now prove that closing
$\CoveredPoly$ under variable assignments is enough to recover from
\cref{karh:thm}. We use the following notation to fix the value of some
indeterminate, if $P(X,Y)$ is a polynomial in $\Rel[X,Y]$, then
$\intro*\restr{P(X,Y)}{X = 1}$ is the polynomial $P(1,Y) \in \Rel[Y]$. More
generally, if $\nu$ is a partial function from $\vec{X}$ to $\Nat$, written
$\nu \colon \vec{X} \topartial \Nat$, the restriction $\restr{P(\vec{X})}{\nu}$
is the polynomial with indeterminates $\vec{Y} \defined \vec{X} - \dom(\nu)$
obtained by fixing the variables of the domain of $\nu$.

@@

\begin{theorem}[label=corrected-version:thm,]
	Let $P \in \Rel[\vec{X}]$.
	The following are equivalent:
	\begin{enumerate}
		\item \label{corrected-1:item} $P \in \CorrectPoly$,
		\item \label{corrected-2:item} $P$ is \kl{represented} by a \kl{$\Nat$-rational series},
		\item \label{corrected-3:item} $P$ is \kl{represented} by a \kl{$\Nat$-polyregular function},
		\item \label{corrected-4:item} $P$ is \kl{represented} by a \kl{star-free $\Nat$-polyregular function},
	\end{enumerate}
	Furthermore, the properties are decidable, and conversions effective.
\end{theorem}

@@

\begin{remark}[restate=undecidable-non-negative:lem,label=undecidable-non-negative:lem]
    \proofref{undecidable-non-negative:lem}
    Checking whether a polynomial
    $P \in \Rel[\vec{X}]$ is \kl{non-negative} is undecidable.
\end{remark}

The proof of \cref{corrected-version:thm} is divided into two parts. First, we
provide in \cref{sec:n-poly-to-poly} a fine combinatorial understanding of what
functions can be computed in $\NPoly$ and $\ZPoly$. This allows us to prove
that \kl{$\Nat$-polyregular functions} are in $\CorrectPoly$
(\cref{n-rat-correct:lem}). Then, in \cref{sec:poly-to-n-poly} we will show how
to compute polynomials in $\CorrectPoly$ using $\NPoly$
(\cref{lem:correct-to-n-rat}). Then, we will next generalize
\cref{corrected-version:thm} to polynomials in $\Rat[\vec{X}]$ in
\cref{sec:rel-to-rat}.

@@

\AP To address our first goal, let us define that a function $q$ is a
\intro{(noncommutative output) polynomial function} from $\Nat^p$ to $\Sigma^*$
whenever there exists words $\alpha_0, \dots, \alpha_p \in \Sigma^*$, and words
$u_1, \dots, u_p \in \Sigma^*$, such that $q(X_1, \dots, X_p) = \alpha_0
\prod_{i = 1}^p u_i^{X_i} \alpha_i$. That is, $q$ is syntactically defined by a
non-commutative monomial over the monoid $\Sigma^*$. \kl{Polynomial functions}
are \kl{commutative} \kl{polyregular functions}.

@@

Let us highlight that the multiplicative factor $\omega$ is necessary in
\cref{n-poly-combinatorics:lem}. Indeed, the function $f \colon \set{a}^* \to
\Nat$ defined as $0$ when the input is of odd length and $1$ when the input is
of even length is \kl{$\Nat$-polyregular}, but $f(a^X)$ is not a polynomial. We
can trade off this multiplicative factor for a constant term addition under the
extra assumption that the function is \kl{star-free polyregular}, as described
in the following \cref{n-sf-combinatorics:lem}. This lemma is not immediately
of use, but is crucial for the upcoming characterization of $\Nat$-rational
polynomials in \cref{decide-rat-poly-npoly:cor}, which in turn is a key
ingredient of our main \cref{zsf-npoly-nsf:thm}.

@@

\begin{corollary}
	\label{n-rat-correct:lem}
	Let $P \in \Rel[X_1, \dots, X_p]$ be an \kl{$\Nat$-rational polynomial}.
	Then,
	$P \in \CorrectPoly$.
\end{corollary}
\begin{proof}
    Let $f$ be a \kl{commutative} \kl{$\Nat$-rational series}
	with domain defined as $\Sigma \defined \set{a_1, \dots, a_p}$
	that \kl{represents} $P$.
    Because $f$ has \kl{polynomial growth},
    $f \in \NPoly$ 
    (\cref{polyregular-polynomial-growth:lemma}).
	Using \cref{n-poly-combinatorics:lem},
	there exists a number $\omega \in \Nat_{\geq 1}$
	and \kl{natural binomial function} $Q$
	such that
	for all $n_1, \dots, n_p \geq 1$:
	\begin{equation*}
		f\left(
		\prod_{i = 1}^p a_i^{\omega n_i}
		\right)
		= Q(n_1, \dots, n_p)
		= P(\omega n_1, \dots, \omega n_p)
		\quad .
	\end{equation*}
    For large enough values of $X$, the \kl{binomial function}
	$\binom{X - p}{k}$ coincides with a polynomial whose dominant coefficient
	is $1/k!$ which is \kl{non-negative}.
	We conclude that
	the \kl{maximal monomials} of
	$P(\omega X_1, \dots, \omega X_p)$ are \kl{non-negative},
	and since $\omega \geq 1$, we conclude that
	the \kl{maximal monomials} of $P$ have \kl{non-negative} coefficients.

	For every partial valuation $\nu \colon \vec{X} \topartial \Nat$,
	the polynomial $\restr{P}{\nu}$ continues to be represented
	by a \kl{$\Nat$-polyregular function}, namely
	$f$ partially applied to a word $u$ and using a restricted input alphabet. As a consequence,
	the \kl{maximal monomials} of
	$\restr{P}{\nu}$ are also \kl{non-negative},
	and
	we have proven that $P \in \CorrectPoly$.
\end{proof}

@@

\AP This section is devoted to proving that polynomials in $\CorrectPoly$ can
be \kl{represented} by \kl{star-free $\Nat$-polyregular functions}. The key
lemma of this section is \cref{lem:correct-to-n-rat}, which is proved by
induction on the number of indeterminates of a given polynomial $P$. In order
to prove that result, we use the combinatorial
\cref{derivation-stabilises-correct:lem} that allows us to transform a
polynomial $P \in \CorrectPoly$ into a polynomial in $\Nat[\vec{X}]$ through a
well-chosen translation of the indeterminates. This argument is based on the
notion of \intro{discrete derivative} which is built by translating the domain
of the polynomial. To that end, let us write $\intro*\translate{K}$ for the
\intro{translation function} that maps a polynomial $P \in \Rel[X_1, \dots,
X_n]$ to the polynomial $P(X_1 + K, \dots, X_n + K)$.

@@

@@

While \cref{lem:correct-to-n-rat} provides an effective conversion procedure,
it does not explicitly state that the membership is decidable to keep the proof
clearer. A similar proof scheme can be followed to conclude that membership is
decidable, and even show that elements in $\CorrectPoly$ are, up to suitable
translations, polynomials in $\Nat[\vec{X}]$
(\cref{derivation-translation:lem}). Beware that partial applications are still
needed in this characterization, as \cref{bad-poly-translate:ex} illustrates.

\begin{lemma}[restate=derivation-translation:lem,label=derivation-translation:lem]
	\proofref{derivation-translation:lem}
	Let $P \in \Rel[\vec{X}]$.
	There exists a computable number $K \in \Nat$
	such that the following are equivalent:
	\begin{enumerate}
		\item \label{d-t-correct:item} $P \in \CorrectPoly$,
		\item \label{d-t-transl:item}
		      for
		      all partial functions $\nu \colon \vec{X} \topartial \Nat$,
		      $\translate{K}(\restr{P}{\nu}) \in \Nat[\vec{X}]$,
		\item \label{d-t-transl-fin:item}
		      for all partial functions
		      $\nu \colon \vec{X} \topartial \set{0, \dots, K}$,
		      $\translate{K}(\restr{P}{\nu}) \in \Nat[\vec{X}]$.
	\end{enumerate}
	In particular, the above properties are decidable.
\end{lemma}

\begin{example}
	\label{bad-poly-translate:ex}
	The polynomial $\BadPoly$ is not a
	\kl{$\Nat$-rational polynomial},
	but is \kl{non-negative} and satisfies
	$\translate{10}(\BadPoly) \in \Nat[\vec{X}]$.
\end{example}

We now have all the tools to prove the corrected version of \cref{karh:thm}.

@@

\AP Let us complete our analysis of polynomials \kl{represented} by $\NRat$
or $\ZRat$ by considering polynomials with coefficients in $\Rat$, and
justify that all the combinatorial work has already happened in $\Rel$ and
$\Nat$. A first observation is that from \cref{n-sf-combinatorics:lem}, we know
that the polynomials that can be computed by \kl{star-free $\Nat$-polyregular
functions} are going to coincide (on large enough inputs) with \kl{natural binomial
functions}. For that reason, we introduce the following ``polynomial
counterpart" of a binomial coefficient: given two numbers $\ell,k \in \Nat$,
$\intro*\pbinom{X - \ell}{k}$ defined as $(X - \ell) \cdots (X - \ell - k) /
k!$,\footnote{ In particular, $\pbinom{X - \ell}{k}$ is defined to be $1$ when
$k = 0$, and $X - \ell$ when $k = 1$. } that we call a \intro{binomial
monomial}, and we introduce \intro{natural binomial polynomials} as $\Nat$-linear
combinations of products of \kl{binomial monomials}, i.e., of the shape:
$P(X_1, \dots, X_k) = \sum_{i = 1}^n n_i \prod_{j = 1}^k \pbinom{X_j -
p_{i,j}}{k_{i,j}}$. Similarly, we introduce the class of \intro{integer
binomial polynomials}, which are obtained by $\Rel$-linear combinations of
products of \kl{binomial monomials}.

\AP
These definitions are justified by the classical result of Pólya that
characterizes polynomials $P$ in $\Rat[X]$ that \intro{integer-valued} (i.e., are such
that $P(\Rel) \subseteq \Rel$) as \kl{integer binomial polynomials}
\cite{POLYA1915,CACHA1996}, which are often called \emph{Newton Polynomials}. Note
that this result straightforwardly extends to multiple indeterminates as we
prove in \cref{integer-binomial-polynomial:lem}.

@@

\begin{theorem}
	\label{integer-binomial-polynomials:cor}
	Let $P \in \Rat[\vec{X}]$, and let $\alpha$ be the minimal number in $\Nat$ such that
	$\alpha P \in \Rel[\vec{X}]$. Then, the following properties are equivalent:
	\begin{enumerate}
        \item \label{int-bin-3:item} $P$ is \kl{integer-valued},
        \item \label{int-bin-2s:item} $P$ is \kl{represented} by a \kl{$\Rel$-rational series},
		\item \label{int-bin-2:item} $P$ is \kl{represented} by a \kl{$\Rel$-polyregular function},
		\item \label{int-bin-1:item} $P$ is \kl{represented} by a \kl{star-free $\Rel$-polyregular function},
		\item \label{int-bin-0:item} $P$ is an \kl{integer binomial polynomial}.
	\end{enumerate}
	These properties are furthermore decidable.
\end{theorem}
\begin{proof}
	The implications \cref{int-bin-1:item}
	$\implies$ \cref{int-bin-2:item} $\implies$ 
    \cref{int-bin-2s:item}
    $\implies$ \cref{int-bin-3:item} are obvious.
	Now, \cref{int-bin-3:item} $\implies$ \cref{int-bin-0:item} is a direct
	consequence of \cref{integer-binomial-polynomial:lem}.
	Finally, \cref{int-bin-0:item} $\implies$ \cref{int-bin-1:item} follows from the fact
	that $\pbinom{X - p}{k}$ is \kl{represented} by a \kl{star-free $\Rel$-polyregular function}
	defined by hardcoding the output values (in $\Rel$) when $0 \leq X \leq p$, and
	using a \kl{star-free $\Nat$-polyregular function} when $X > p$
	(\cref{binomial-function-star-free:lem}).
	Because
	$\ZSF$ is closed under products and $\Rel$-linear combinations, we conclude.
\end{proof}

\AP Obtaining an analogue of \cref{integer-binomial-polynomials:cor} for
\kl{$\Nat$-polyregular functions} requires a bit more work, as polynomials in
$\Rat[\vec{X}]$ that are \kl{represented} by $\NPoly$ are not exactly
\kl{natural binomial polynomials}
(see \cref{natural-binomial-polynomial-positive-bad:ex}). To address the issues
raised by the former example, we introduce the notion of \intro{strongly
natural binomial polynomials}, as the polynomials $P \in \Rat[X]$ such that for
all partial valuation $\nu \colon \topartial \Nat$, $\restr{P}{\nu}$ is a
\kl{natural binomial polynomial}, which characterizes the class of polynomials
that are \kl{represented} by $\NPoly$ (\cref{decide-rat-poly-npoly:cor}).

\begin{example}[restate=natural-binomial-polynomial-positive-bad:ex,label=natural-binomial-polynomial-positive-bad:ex]
	%\proofref{natural-binomial-polynomial-positive-bad:ex}
	The polynomial
	$Q(X,Y,Z) \defined \pbinom{X - 4}{1} \pbinom{Y}{1} \pbinom{Z}{1}
		+ 8 \pbinom{Y}{2} + 8 \pbinom{Z}{2} + 4$
	is a \kl{non-negative} \kl{natural binomial polynomial},
	but cannot be computed by a \kl{star-free $\Nat$-polyregular function}.
    Indeed, 
    $Q(0,Y,Z)$ has a negative maximal monomial, hence $Q \not \in \CorrectPoly$,
    and we conclude using 
    \cref{corrected-version:thm}.
\end{example}

\begin{lemma}[restate=polyrec-integer-strong:lem,label=polyrec-integer-strong:lem]
    \proofref{polyrec-integer-strong:lem}
    Let $P \in \Rat[\vec{X}]$ be an \kl{integer-valued} polynomial, 
    and $n \in \Nat_{\geq 1}$ be
    such that $n P \in \Rel[\vec{X}] \cap \CorrectPoly$.
    Then, $P$ is a \kl{strongly natural binomial polynomial}.
\end{lemma}

\begin{theorem}
	\label{decide-rat-poly-npoly:cor}
	Let $P \in \Rat[\vec{X}]$ be a polynomial with \emph{rational}
	coefficients and let $\alpha$ be the smallest number in $\Nat_{\geq 1}$
	such that $\alpha P \in \Rel[\vec{X}]$. Then, 
    the following are equivalent:
	\begin{enumerate}
        \item \label{rat-npoly-4:item} $\alpha P \in \CorrectPoly$ 
            and $P$ is \kl{integer-valued},
        \item \label{rat-npoly-0:item} $P$ is \kl{represented} by a \kl{$\Nat$-rational series},
		\item \label{rat-npoly-1:item} $P$ is \kl{represented} by a \kl{$\Nat$-polyregular function},
		\item \label{rat-npoly-2:item} $P$ is \kl{represented} by a \kl{star-free $\Nat$-polyregular function},
		\item \label{rat-npoly-3:item}
		      $P$ is a \kl{strongly natural binomial polynomial}.
	\end{enumerate}
	In particular, the properties are decidable.
\end{theorem}
\begin{proof}
    Let us first remark that $\NPoly \subseteq \NRat$,
    and that if $P$ is \kl{represented} by a function $f \in \NRat$,
    then said function has \kl{polynomial growth}, and in particular
    $f \in \NPoly$ thanks to \cref{polyregular-polynomial-growth:lemma}.
    As a consequence, \cref{rat-npoly-0:item} $\iff$ \cref{rat-npoly-1:item}.
    For the implication \cref{rat-npoly-1:item} $\implies$
    \cref{rat-npoly-4:item}, we obtain $\alpha P \in \CorrectPoly$ via
    \cref{corrected-version:thm} by remarking that the \kl{star-free
    $\Nat$-polyregular functions} have output in $\Nat$ and are closed under
    multiplication by a constant $\alpha \in \Nat$. The fact that $P$ is
    \kl{integer-valued} follows from
    \cref{integer-binomial-polynomials:cor} and the fact that
    $\NSF \subseteq \ZPoly$.
    The implication \cref{rat-npoly-4:item} $\implies$
    \cref{rat-npoly-3:item} is obtained thanks to \cref{polyrec-integer-strong:lem}.
